Neural Models for Target-Based Computer-Assisted Musical Orchestration: A Preliminary Study

### 2021-06-25 csmc_mume_2020/paper/JCMS_2021/paper.tex
Revise conclusion

Partial feature dropout was performed on the feature matrix itself of input samples, the Mel spectrogram. We chose random columns and rows of the matrix to zero out. For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out. This method of data augmentation aimed to be more resilient to the possible variations in the recording of the instruments.

@@

\begin{table*}
	\centering    
  \resizebox{\columnwidth}{!}{\begin{tabular}{|c|c|c|c|c|c|c|c|c|c}
     \hline
      \textbf{Model} & \textbf{Ob + Bn} & \textbf{Bn} & \textbf{Bass cl.} & \textbf{Bell 1} & \textbf{Bell 2} & \textbf{Multiph. 1} & \textbf{Car horn} & \textbf{Boat} \\
      \hline
      CNN with LSTM & 0.17 & 0.28 & 0.70 & 0.55 & 0.26 & \textbf{1.10} & 0.68 & \textbf{1.12} \\
      \hline
      ResNet & 0.34 & 0.50 & 0.48 & 0.59 & 0.45 & 0.90 & 0.49 & \textbf{1.16} \\
      \hline
      \hline
      \textbf{Model} & \textbf{Wind harp} & \textbf{Chord 1} & \textbf{Multiph. 2} & \textbf{Chord 2} & \textbf{Gong} & \textbf{Scream} & \textbf{Brass} & \textbf{Average}  \\
      \hline
      CNN with LSTM & 0.55 & 0.79 & 0.70 & 0.57 & 0.73 & \textbf{1.14} & 0.79 & 0.71\\
      \hline
      ResNet  & 0.61 & 0.86 & 0.51 & 0.37 & 0.71 & \textbf{1.03} & \textbf{1.05} & 0.66 \\
      \hline
    \end{tabular}}
    \caption{Quantitative comparison of orchestrations as ratios to Orchidea. Eqn. \ref{distance} was used to compute distances between orchestrations and targets. What is shown is the ratio between the distance of Orchidea's solution to the target and our solution's distance to the same target. A value less than 1 means that our model performed worse (i.e. had a larger distance), and a value greater than 1 means our model performed better than Orchidea. The last column shows the ratio of the average distances for the model across all targets.}\label{orch_eval}
\end{table*}

@@

Orchidea implements many advanced features that are not supported by our models. For example, it is able to apply symbolic constraints to the search, hence allowing only specific instrumental combinations or playing styles. It is also able to reduce the search space by applying harmonic analysis on the target sound. The dominant harmonic partials of the target are identified, and the search space is limited to only include samples of those pitches. For example, if the target is a recording of an instrument playing a C4, then the partials identified may be C4, C5, G5, and E6. The model would then only consider samples of these pitches to be used in the solution. This leads to a solution whose harmonics are much closer to the target, which is an important part of aural similarity. Orchidea's solutions, moreover, can use any number of instruments included in the orchestra specified by the user, thus having variable-sized orchestrations from a single instrument to the whole set (this property is called \emph{sparsity} of the solution).

@@

We calculated the distance between the target samples and our orchestrated solutions. We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions. A comparison of these results is in Table \ref{orch_eval}. While our model is not able to outperform Orchidea, it shows consistent results. We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets. Quantitatively, CNN performs better with an average ratio of $0.71$ compared to ResNet with an average ratio of $0.66$. Both models outperform Orchidea on three out of fifteen of the targets. Through our qualitative inspection, we find that CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.

@@

@@

>>>>

Partial feature dropout was performed on the feature matrix itself of input samples, the Mel spectrogram. We chose random columns and rows of the matrix to zero out. For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out.

@@

\begin{table*}
	\centering    
  \resizebox{\columnwidth}{!}{\begin{tabular}{|c|c|c|c|c|c|c|c|c|c}
     \hline
      \textbf{Model} & \textbf{Ob + Bn} & \textbf{Bn} & \textbf{Bass cl.} & \textbf{Bell 1} & \textbf{Bell 2} & \textbf{Multiph. 1} & \textbf{Car horn} & \textbf{Boat} \\
      \hline
      CNN with LSTM & 0.17 & 0.28 & 0.70 & 0.55 & 0.26 & \textbf{1.10} & 0.68 & \textbf{1.12} \\
      \hline
      ResNet & 0.34 & 0.50 & 0.48 & 0.59 & 0.45 & 0.90 & 0.49 & \textbf{1.16} \\
      \hline
      \hline
      \textbf{Model} & \textbf{Wind harp} & \textbf{Chord 1} & \textbf{Multiph. 2} & \textbf{Chord 2} & \textbf{Gong} & \textbf{Scream} & \textbf{Brass} & \textbf{Average}  \\
      \hline
      CNN with LSTM & 0.55 & 0.79 & 0.70 & 0.57 & 0.73 & \textbf{1.14} & 0.79 & 0.71\\
      \hline
      ResNet  & 0.61 & 0.86 & 0.51 & 0.37 & 0.71 & \textbf{1.03} & \textbf{1.05} & 0.66 \\
      \hline
    \end{tabular}}
    \caption{Quantitative comparison of orchestrations as ratios to Orchidea. Eqn.~\ref{distance} was used to compute distances between orchestrations and targets. What is shown is the ratio between the distance of Orchidea's solution to the target and our solution's distance to the same target. A value less than 1 means that our model performed worse (i.e. had a larger distance), and a value greater than 1 means our model performed better than Orchidea. The last column shows the ratio of the average distances for the model across all targets.}\label{orch_eval}
\end{table*}

@@

Orchidea implements many advanced features that are not supported by our models. For example, it is able to apply symbolic constraints to the search, hence allowing only specific instrumental combinations or playing styles. It is also able to reduce the search space by applying harmonic analysis on the target sound. The dominant harmonic partials of the target are identified, and the search space is limited to only include samples of those pitches. For example, if the target is a recording of an instrument playing a C4, then the partials identified may be C4, C5, G5, and E6. The model would then only consider samples of these pitches to be used in the solution. This leads to a solution whose harmonics are much closer to the target, which is an important part of aural similarity. Orchidea's solutions, moreover, can use any number of instruments included in the orchestra specified by the user, thus having variable-sized orchestrations from a single instrument to the whole set (this property is called the \emph{sparsity} of the solution).

@@

We calculated the distance between the target samples and our orchestrated solutions. We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions. A comparison of these results is in Table~\ref{orch_eval}.

@@

While our models are not able to outperform Orchidea, they shows consistent results. During training, CNN and ResNet both perform well, with ResNet consistently achieving higher training accuracies than CNN. For orchestration, CNN performs quantitatively better with an average ratio of $0.71$ compared to ResNet with an average ratio of $0.66$ (see Table \ref{orch_eval}). Both models outperform Orchidea on three out of fifteen of the targets. Through our qualitative inspection, we find that CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.

@@

Finally, the OrchideaSOL dataset could replace TinySOL as the dataset used. OrchideaSOL could improve the system by adding extended playing techniques to the data, which can allow noisier targets to be better orchestrated. OrchideaSOL contains 13,265 samples from 14 instruments, almost ten times as many samples as TinySOL~\cite{Cella2020b}. This would greatly increase the number of classes, and therefore we would need to train the models on more samples and possibly even use deeper models.
